---
title: "GVAR_mplus.qmd"
format: html
editor: source
author: Björn S. Siepe
---

# Preparation
Load relevant packages
```{r}
library(graphicalVAR)
library(here)
library(tidyverse)
library(MplusAutomation)
set.seed(35037)
```

Create dataset to feed into MPlus:
```{r}
# Additionally, use data from our preprint
l_graphs <- readRDS(here("data", "l_graphs.rds"))

# Choose graph 6
gvar_mod <- l_graphs$graph6

# Add partial correlations
gvar_mod$PCC <- -cov2cor(gvar_mod$kappa)
diag(gvar_mod$PCC) <- 0

# Simulate data
gvar_dat <- as.data.frame(graphicalVAR::graphicalVARsim(nTime = 200, 
                                          beta = gvar_mod$beta[1:4, 1:4],
                                          kappa = gvar_mod$kappa[1:4, 1:4]))
gvar_mod$beta[1:4, 1:4]

```


# Prepare MPlus
```{r}
# Prepare MPlus data via MPlusautomation
# keep all variables
prepareMplusData(
  gvar_dat,
  filename = here("scripts/mplus_gvar_example.dat")) # scripts folder for Mplus
```

Create MPlus input text file
```{r}
# Create MPlus input text file
mplus_model <- " 
DATA:	    FILE = mplus_gvar_example.dat;    ! data file (should be in same folder)

VARIABLE:	NAMES = V1 V2 V3 V4;          ! providing names to the variables 
            USEVARIABLES = V1 V2 V3 V4;   ! select variables for the analysis
	        LAGGED = V1(1) V2(1) V3(1) V4(1);  ! creating first-order
                                    ! lagged observed variables                                    
            MISSING = *;            ! missing value code

ANALYSIS:	ESTIMATOR = BAYES;      ! set estimator (must be Bayes for DSEM) 
	        PROCESSORS = 8;         ! using 2 processors
	        BITERATIONS = (4000);   ! choose number of iterations;
                                    ! minimum is now 2000; will be more if 
                                    ! the convergence criterion indicates
                                    ! convergence was not reached

MODEL:	    V1 ON V1&1 V2&1 V3&1 V4&1;        
            V2 ON V1&1 V2&1 V3&1 V4&1;
            V3 ON V1&1 V2&1 V3&1 V4&1;
            V4 ON V1&1 V2&1 V3&1 V4&1;

OUTPUT:	    TECH1 TECH8;            ! asking additional output
SAVEDATA: BPARAMETERS IS additional_mplus.dat; ! saving posterior samples
PLOT:	    TYPE = PLOT3;           ! asking for plots
"
# Write MPlus input file
writeLines(mplus_model, here("scripts/mplus_gvar_example.inp"))

```

Now execute the model
```{r}
# Now execute the model
# use the syntax above for the relevant model
# mplusModeler(TITLE = "GVAR Example",
#              DATA = "mplus_gvar_example.dat",
#              OUTPUT = "mplus_gvar_output.out",
#              MODEL = "V1 ON V1&1 V2&1 V3&1 V4&1;
#                       y2 ON V1&1 V2&1 V3&1 V4&1;
#                       y3 ON V1&1 V2&1 V3&1 V4&1;
#                       y4 ON V1&1 V2&1 V3&1 V4&1;"
#              RUNS = 1,
#              PROCESSORS = 8,
#              BURNIN = 2000,
#              SAMPLES = 2000,
#              CHAINS = 4,
#              SEED = 35037,
#              )


runModels(here("scripts/mplus_gvar_example.inp"), 
                       showOutput=TRUE)
```


Extract posterior samples
```{r}
# Extract posterior samples
mplus_res <- readModels(here("scripts/mplus_gvar_example.out"))

mplus_samples <- mplus_res$bparameters$valid_draw

# convert the list mplus_samples to a dataframe
mplus_samples <- do.call(rbind, mplus_samples)
mplus_samples <- as.data.frame(mplus_samples)

names(mplus_samples)

# Extract every column that contains an ".ON" in the name
# these are the regression coefficients
samples_beta <- mplus_samples %>% 
  dplyr::select(contains(".ON")) 
  

# extract innovation covariance
samples_theta <- mplus_samples %>% 
  dplyr::select(contains("with"))

# extract diagonal columns of covariance matrix
selected_columns <- names(mplus_samples) %>% 
  str_detect("\\d+_V\\d$")
samples_diag <- mplus_samples %>% 
  dplyr::select(which(selected_columns))


```

Convert posterior samples to matrix format for beta is rather straightforward:
```{r}
# Convert posterior samples to matrix format
# beta
split_beta <- split(samples_beta, 1:nrow(samples_beta))
beta_samples <- lapply(split_beta, function(x) {
  matrix(x, nrow = 4, ncol = 4, byrow = TRUE)
})
# convert this list to an array
beta_samples <- array(unlist(beta_samples), dim = c(4, 4, length(beta_samples)))






```

More complicated for theta:
```{r}
# Function to extract covariance values and create covariance matrix
create_cov_mat <- function(dat, n_row, n_col) {
  
  cov_matrix <- matrix(NA, nrow = n_row, ncol = n_col)
  diag(cov_matrix) <- 0
  for (i in 1:ncol(dat)) {
    col_name <- names(dat)[i]
    values <- as.vector(strsplit(sub(".*_", "", col_name), "\\."))  # Extract everything after the underscore
    
    # Extract row and column indices from the column name
    row_index <- as.numeric(gsub("V", "", values[[1]][1]))
    col_index <- as.numeric(gsub("V", "", values[[1]][3]))
    
    # Assign the value to the corresponding position in the covariance matrix
    cov_matrix[row_index, col_index] <- as.numeric(dat[,i])
    cov_matrix[col_index, row_index] <- as.numeric(dat[,i])  # Covariance matrix is symmetric
  }
  
  return(cov_matrix)
}

# theta
split_theta <- split(samples_theta, 1:nrow(samples_theta))
theta_samples <- lapply(split_theta, function(x) {
  create_cov_mat(x, 4, 4)
})

# fill each diagonal with the variances
for (i in 1:length(theta_samples)) {
  diag(theta_samples[[i]]) <- as.numeric(samples_diag[i,])
}

# Convert to array
theta_samples <- array(unlist(theta_samples), dim = c(4, 4, length(theta_samples)))



```

Convert to partial correlations
```{r}
pcor_samples <- lapply(theta_samples, function(x){
  -cov2cor(x)
})
# convert to array
pcor_samples <- array(unlist(pcor_samples), dim = c(4, 4, length(pcor_samples)))

```


Convert this all to a function: 
The only issue is that this need variable naming with "V"...
```{r}
library(tidyverse)

convert_mplus_samples <- function(mplus_output_file) {
  # browser()
  # Extract posterior samples
  mplus_res <- readModels(mplus_output_file)
  mplus_samples <- as.data.frame(do.call(rbind, mplus_res$bparameters$valid_draw))

  # Extract columns
  samples_beta <- mplus_samples %>% 
    dplyr::select(contains(".ON"))
  samples_theta <- mplus_samples %>% 
    dplyr::select(contains("with"))
  selected_columns <- names(mplus_samples) %>% 
    str_detect("\\d+_V\\d$")
  samples_diag <- mplus_samples %>% 
    dplyr::select(which(selected_columns))
  nvar <- sum(selected_columns)  # Number of variables

  # Convert posterior samples to matrix format
  beta_samples <- lapply(split_beta, function(x) {
    matrix(x, nrow = 4, ncol = 4, byrow = TRUE)
  })
  # convert this list to an array
  beta_samples <- array(unlist(beta_samples), 
                        dim = c(4, 4, length(beta_samples)))
  split_theta <- split(samples_theta, 1:nrow(samples_theta))
  

  # Function to extract covariance values and create covariance matrix
  create_cov_mat <- function(dat, diag_values) {
    cov_matrix <- matrix(NA, nrow = nvar, ncol = nvar)
    diag(cov_matrix) <- 0
    for (i in 1:ncol(dat)) {
      col_name <- names(dat)[i]
      values <- as.vector(strsplit(sub(".*_", "", col_name), "\\."))

      # Extract row and column indices from the column name
      row_index <- as.numeric(gsub("V", "", values[[1]][1]))
      col_index <- as.numeric(gsub("V", "", values[[1]][3]))

      # Assign the value to the corresponding position in the covariance matrix
      cov_matrix[row_index, col_index] <- as.numeric(dat[, i])
      cov_matrix[col_index, row_index] <- as.numeric(dat[, i])  # Covariance matrix is symmetric
    }
    diag(cov_matrix) <- as.numeric(diag_values)
    return(cov_matrix)
  }

  # this is a very ugly mix of a loop and lapply, sorry...
  theta_samples <- simplify2array(lapply(seq_along(split_theta), function(i) {
      create_cov_mat(split_theta[[i]], samples_diag[i,])
      }))


  # Convert to partial correlations
  pcor_samples <- array(NA, dim = dim(theta_samples))
  for(i in 1:dim(theta_samples)[3]) {
    tmp <- -cov2cor(theta_samples[,,i])
    diag(tmp) <- 0
    pcor_samples[, , i] <- tmp
  }

  return(list(beta_samples = beta_samples, theta_samples = theta_samples, pcor_samples = pcor_samples))
}

```

Try the function:
```{r}
test <- convert_mplus_samples(here("scripts/mplus_gvar_example.out"))

```






